---
title: "Interaction analysis"
author: "Lars Mølgaard Saxhaug"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
library(tidybayes)
library(here)

options(contrasts=c('contr.treatment','contr.treatment'))
```

Disclaimer: This is NOT a proper reanalysis (only loosely based on [this](https://www.poverty-action.org/publication/impact-community-masking-covid-19-cluster-randomized-trial-bangladesh) trial), but a proof of concept of analysis of interaction using bayesian methods.

```{r data_import}
df <- tribble(
 ~age,      ~n,     ~intervention_prevalence_ratio , ~control_prevalence,
 "=< 40",  147954,   0.977,                            0.0055,
 "40-50",  36002,    1.003,                            0.0095,
 "50-60",  24282,    0.768,                            0.0108,
 ">60",    28102,    0.669,                            0.0103   # n reduced to from 28103 for simplicity
   
) %>% 
  mutate(control_n=n/2,
         intervention_n=n/2,
         control_events=round(control_n*control_prevalence),
         intervention_events=round(intervention_n*control_prevalence*intervention_prevalence_ratio)) %>% 
  select(age,starts_with(c("control","intervention")),-contains(c("ratio","prevalence"))) %>% 
  pivot_longer(cols = starts_with(c("control","intervention")),names_to=c("group","variable"),names_sep = "_") %>% 
  pivot_wider(names_from = "variable",values_from = "value") %>% 
  mutate(across(age,~ordered(.x,levels=c("=< 40","40-50","50-60",">60"))),across(group,as.factor)) 
```

Frequentist test of interaction

```{r freq_fit1}
# treatment effect only
fmod1 <- glm(cbind(events,n-events)~group,family = binomial,data = df)
summary(fmod1)

```

```{r freq_fit2}
# treatment and age effect, but no interaction
fmod2 <- glm(cbind(events,n-events)~group+age,family = binomial,data = df)
summary(fmod2)
```

```{r freq_fit3}
# treatment, age and interaction effect
fmod3 <- glm(cbind(events,n-events)~group*age,family = binomial,data = df)
summary(fmod3)
```

Model that includes interaction term is a significantly better fit to the data than the base model, but a smaller incremental improvement than including age specific intercepts

```{r freq_compare}
anova(fmod1,fmod2,fmod3,test = "LRT")

```

# Bayesian analysis

Priors:

```{r priors}
prior_base <- prior(normal(0,1),class="b")
prior_mo <- prior(normal(0,1),class="b")+
            prior(dirichlet(1),class="simo",coef="moage1")
prior_mo_interaction <-prior(normal(0,1),class="b")+
            prior(dirichlet(1),class="simo",coef="moage1")+
            prior(dirichlet(1), class= "simo", coef="moage:groupintervention1")
prior_onesided <- prior(normal(0,1),class="b",ub=0)+
            prior(normal(0,1),class="b", coef="moage")+
            prior(dirichlet(1),class="simo",coef="moage1")+
            prior(dirichlet(1), class= "simo", coef="moage:groupintervention1")
```

Model 1: simple for overall effect without interaction with age

```{r fit_1a}
# simple model for overall effect
mod1_a <- brm(events|trials(n)~group,family = binomial,data = df,file = here("fits","mod1.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior = prior_base)
mod1_a
plot(mod1_a)
```

Model 1 b: Added age specific intercept (age specific infection risk but not interaction)

```{r fit_1b}
# added age specific intercept
mod1_b <- brm(events|trials(n)~group+age,family = binomial,data = df,file = here("fits","mod1_b.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior = prior_base)
mod1_b
plot(mod1_b)
```

Model 2 a: Added interaction between categorical age and intervention

```{r fit_2a}
# model with added interaction with categorical age
mod2_a <- brm(events|trials(n)~group*age,family = binomial,data = df,file = here("fits","mod2.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior = prior_base)
mod2_a
plot(mod2_a)
```

Model 3_a: Age as monotonic effect (on infection risk but no interaction)

Brms implements monotonic effects for ordinal variables as described [here](https://psyarxiv.com/9qkhj), "bsp_mo" parameters describe the difference in response between minimum and maximum, simplex parameters subdivide the effect on the ordinal levels

```{r fit_3a}
# model with added  age as monotonic effect
mod3_a <- brm(events|trials(n)~group+mo(age),family = binomial,data = df,file = here("fits","mod3.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior = prior_mo)
mod3_a
plot(mod3_a)
```

Model 3_b: Interaction between monotonic age effect and intervention

```{r fit_3b}
# model with added interaction with age as monotonic effect
mod3_b <- brm(events|trials(n)~group*mo(age),family = binomial,data = df,file = here("fits","mod3_c.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior = prior_mo )
mod3_b
plot(mod3_b)
get_variables(mod3_b)
pairs(mod3_b)
mod3_b %>% 
  spread_draws(`b_groupintervention`,`bsp_moage:groupintervention` ) %>% 
  ggplot(aes(x=b_groupintervention,y=`bsp_moage:groupintervention`))+
  geom_point(alpha=0.2)+
  theme_tidybayes()+
  labs(title = "Scatterplot of posterior values of overall effect\n(b_groupintervention) with interaction effect (bsp)",caption = "@load_dependent")
```

Model 4: Prior ruling out negative effect of intervention

```{r fit_4}
# as model 3_a but with prior ruling out negative effect
mod4 <- brm(events|trials(n)~group*mo(age),family = binomial,data = df,file = here("fits","mod3_b.rds"),file_refit = "on_change",save_pars = save_pars(all = TRUE),prior=prior_onesided)
mod4
plot(mod4)
```

Setting up model comparisions using "leave one out cross validation"

```{r model_comparison}
mod1_a <- mod1_a %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)
mod1_b <- mod1_b %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)
mod2_a <- mod2_a %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)
mod3_a <- mod3_a %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)
mod3_b <- mod3_b %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)
mod4 <- mod4 %>% add_criterion(c("loo","waic"), moment_match=TRUE,reloo=TRUE)


l <- loo_compare(mod1_a,mod1_b,mod2_a,mod3_a,mod3_b)

cbind(loo_diff = l[, 1] * -2,
      se       = l[, 2] *  2) %>% 
  knitr::kable()
```

```{r model_comparison_weights}
model_weights(mod1_a,mod1_b,mod2_a,mod3_a,mod3_b,weights = "loo") %>% round(digits = 6) %>% 
  knitr::kable()

```

```{r}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod2_a,scale="response") %>% 
  compare_levels(.value,by = group,fun=`/`) %>%
  median_hdi(.value) %>% 
  knitr::kable()
```

```{r}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod3_b,scale="response") %>% 
  compare_levels(.value,by = group,fun=`/`) %>%
  compare_levels(.value,by=age,comparison="pairwise") %>% 
  median_hdi() %>% 
  knitr::kable()
```

```{r plot_1a}
df %>% 
  modelr::data_grid(group,n=1) %>% 
  add_linpred_draws(mod1_a,scale="response") %>% 
  compare_levels(.value,by = group,fun=`/`)%>% 
  ggplot(aes(x=.value,fill=after_stat(x>1)))+
  stat_slab(position="dodge")+
  stat_dotsinterval(side="bottom", quantiles=100,position ="dodge",fill="grey")+
  scale_fill_manual(values=c("lightblue","pink"))+
  scale_x_continuous(name="Relative risk")+
  scale_y_continuous(name = "")+
  theme_tidybayes()+
  theme(legend.position = "none",axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  labs(title = "Model 1: average effect regardless of age",
       caption = "@load_dependent")
```

```{r plot_interaction_1b}
df %>% 
  modelr::data_grid(group,age,n=1) %>% 
  add_linpred_draws(mod1_b,scale="response") %>% 
  compare_levels(.value,by = group,fun=`/`)%>% 
 ggplot(aes(x=.value,y=age,fill=after_stat(x>1)))+
  stat_slab(position="dodge",scale=0.5)+
  stat_dotsinterval(side="bottom", quantiles=100,position ="dodge",scale=0.5,fill="grey")+
  scale_fill_manual(values=c("lightblue","pink"))+
  scale_y_discrete(name = "Age group")+
  scale_x_continuous(name="Relative risk")+
  theme_tidybayes()+
  theme(legend.position = "none")+
  labs(title = "Model 1 b: age specific infection risk only ",
caption = "@load_dependent")
```

```{r plot_interaction_2a}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod2_a,scale="linear") %>% 
  compare_levels(.value,by = group,fun = `/`,comparison = rlang::exprs(control / intervention  )) %>%
  ggplot(aes(x=.value,y=age,fill=after_stat(ifelse(x>1,"over","under"))))+
  stat_slab(position="dodge",scale=0.6)+
  stat_dotsinterval(side="bottom", quantiles=100,position ="dodge",scale=0.6,fill="grey")+
  scale_fill_manual(values=c("pink","lightblue"))+
  scale_y_discrete(name = "Age group")+
  scale_x_continuous(name="Relative risk")+
  theme_tidybayes()+
  theme(legend.position = "none")+
  labs(title = "Model 2: age specific infection risk \nand interaction with intervention ",
       caption = "@load_dependent")
```

```{r plot_interaction_3a}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod3_a,scale="response") %>% 
  compare_levels(.value,by = group,fun = `/`) %>%
  ggplot(aes(x=.value,y=age,fill=after_stat(ifelse(x>1,"over","under"))))+ 
  stat_slab(position="dodge",scale=0.6)+
  stat_dotsinterval(side="bottom", quantiles=100,position ="dodge",scale=0.6,fill="grey")+
  scale_fill_manual(values=c("pink","lightblue"))+
  scale_y_discrete(name = "Age group")+
  scale_x_continuous(name="Relative risk")+
  theme_tidybayes()+
  theme(legend.position = "none")+
  labs(title = "Model 3 a: monontonic effect of age on infection risk, without interaction term",
       caption = "@load_dependent")
```

```{r plot_interaction_2b}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod3_b,scale="response") %>% 
  compare_levels(.value,by = group,fun = `/`) %>%
  ggplot(aes(x=.value,y=age,fill=after_stat(x>1)))+ 
  stat_slab(position="dodge",scale=0.6)+
  stat_dotsinterval(side="bottom", quantiles=100,position ="dodge",scale=0.6,fill="grey")+
  scale_fill_manual(values=c("lightblue","pink"))+
  scale_y_discrete(name = "Age group")+
  scale_x_continuous(name="Relative risk")+
  theme_tidybayes()+
  theme(legend.position = "none")+
  labs(title = "Model 3 b: monontonic effect of age on infection risk \nincluding interaction with intervention effect",
       caption = "@load_dependent")

```

```{r plot_interaction_4}
df %>% 
  modelr::data_grid(age,group,n=1) %>% 
  add_linpred_draws(mod4,scale="response") %>% 
  compare_levels(.value,by = group,fun = `/`) %>%
  mutate(or=exp(.value)) %>% 
  ggplot(aes(x=.value,y=age,fill=after_stat(ifelse(x>1,"over","under"))))+
  stat_halfeye()+
  scale_fill_manual(values=c("pink","lightblue"))+
  theme_tidybayes()+
  theme(legend.position = "none")+
  labs(title = "Model 4 (prior ruling out neg effect)",
       caption = "@load_dependent")

```
